---
title: "lecture 5"
author: "Natalie Benoit"
date: "2/24/2021"
output: html_document
---

# install packages
```{r}
library(tidyverse)
library(nycflights13)
library(PNWColors)
```

# looking at the data 
```{r}
head(airports)
head(weather)
head(planes)
head(airlines)
```

### PART 1

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?
# I would need to combine the plane and flight tables

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?
# It is related though weather, flights, and airports. The origiin is the same. 

weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?
# Weather, flights, athrough destination to complete this relationship. 


### PART 2
Mutating joins
```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
flights2

flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")

flights2 %>%
  select(-origin, -dest) %>% 
  left_join(airlines, by = "carrier")

```

Understanding joins
```{r}
x <- tribble(
  ~key, ~val_x,
     1, "x1",
     2, "x2",
     3, "x3"
)
y <- tribble(
  ~key, ~val_y,
     1, "y1",
     2, "y2",
     4, "y3"
)

x %>% 
  inner_join(y, by = "key")
```


13.4.6 (1-3)
```{r}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat, color = tz)) +
    borders("state") +
    geom_point() +
    coord_quickmap() 

airport_org_dest <- airports %>%
  select(faa, lat, lon)

flights %>%
  select(year:day, hour, origin, dest) %>%
  left_join(
    airport_org_dest,
    by = c("origin" = "faa")
  ) %>%
  left_join(
    airport_org_dest,
    by = c("dest" = "faa")
  )



plane_cohorts <- inner_join(flights,
  select(planes, tailnum, plane_year = year),
  by = "tailnum") %>%
  mutate(age = year - plane_year) %>%
  filter(!is.na(age)) %>%
  mutate(age = if_else(age > 21, 21L, age)) %>%
  group_by(age) %>%
  summarise(
    dep_delay_mean = mean(dep_delay, na.rm = TRUE),
    dep_delay_sd = sd(dep_delay, na.rm = TRUE),
       n_arr_delay = sum(!is.na(arr_delay)),
    n_dep_delay = sum(!is.na(dep_delay))
  )
# I chose 21 as the age as this indicates the plane was made in the 90s. 
    
```


### PART 3
13.5 (2 & 5)
```{r}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)
top_dest

flights %>% 
  filter(dest %in% top_dest$dest)

flights %>% 
  semi_join(top_dest)

flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)

```


```{r}
library(fueleconomy)

fueleconomy::vehicles %>%
  semi_join(fueleconomy::common, by = c("make", "model"))


anti_join(flights, airports, by = c("dest" = "faa"))
# This tells me the destinations that do not align with each airport code. This could mean the airport destination list is not very extensive. 

anti_join(airports, flights, by = c("faa" = "dest"))
# This tells me the airport codes that did not align with a destination. As you can see,this is a much smaller table indicating that these airports were not as common destinations. 
```



