---
title: "Lecture6"
author: "Natalie Benoit"
date: "3/2/2021"
output: html_document
---
# Set up
```{r setup, include=FALSE}
library(gapminder)
library(tidyverse)
library(dplyr)
library(stringr)
gDat <- gapminder::gapminder
str(gDat)
```


# Part 1
```{r}
grep('\'', levels(gDat$country), value = TRUE)
```

\': single quote. You don't need to escape single quote inside a double-quoted string, so we can also use "'" in the previous example.
\": double quote. Similarly, double quotes can be used inside a single-quoted string, i.e. '"'.
\n: newline.
\r: carriage return.
\t: tab character.

cat() and print() to handle escape sequences differently, if you want to print a string out with these sequences interpreted, use cat().
```{r}
print("a\nb")
cat("a\nb")
```

Quantifiers
Quantifiers specify how many repetitions of the pattern.

*: matches at least 0 times.
+: matches at least 1 times.
?: matches at most 1 times.
{n}: matches exactly n times.
{n,}: matches at least n times.
{n,m}: matches between n and m times.

```{r}
(strings <- c("a", "ab", "acb", "accb", "acccb", "accccb"))
grep("ac*b", strings, value = TRUE)
grep("ac+b", strings, value = TRUE)
grep("ac?b", strings, value = TRUE)
grep("ac{2}b", strings, value = TRUE)
grep("ac{2,}b", strings, value = TRUE)
grep("ac{2,3}b", strings, value = TRUE)
```


# Exercise 1: 
line of code to find all countried with ee
```{r}
grep("e{2}", levels(gDat$country), value = TRUE)
```

# Part 2

Position of pattern within the string
^: matches the start of the string.
$: matches the end of the string.
\b: matches the empty string at either edge of a word. Don't confuse it with ^ $ which marks the edge of a string.
\B: matches the empty string provided it is not at an edge of a word.

```{r}
(strings <- c("abcd", "cdab", "cabd", "c abd"))
grep("ab", strings, value = TRUE)
grep("^ab", strings, value = TRUE)
grep("ab$", strings, value = TRUE)
grep("\\bab", strings, value = TRUE)
```

# Exercise 2:
write a code to find all files in main folder of repository
```{r}
path_towd <- getwd() # to get working directory
dir(path_towd) #see file names
dir(path_towd, pattern = ".R") # select for .R only -> does include .Rproj. Could I also do #.Rmd? 
```

# Part 3

Operators

.: matches any single character, as shown in the first example.
[...]: a character list, matches any one of the characters inside the square brackets. We can also use - inside the brackets to specify a range of characters.
[^...]: an inverted character list, similar to [...], but matches any characters except those inside the square brackets.
\: suppress the special meaning of metacharacters in regular expression, i.e. $ * + . ? [ ] ^ { } | ( ) \, similar to its usage in escape sequences. Since \ itself needs to be escaped in R, we need to escape these metacharacters with double backslash like \\$.
|: an "or" operator, matches patterns on either side of the |.
(...): grouping in regular expressions. This allows you to retrieve the bits that matched various parts of your regular expression so you can alter them or use them for building up a new string. Each group can than be refer using \\N, with N being the No. of (...) used. This is called backreference.

```{r}
(strings <- c("^ab", "ab", "abc", "abd", "abe", "ab 12"))
grep("ab.", strings, value = TRUE)
grep("ab[c-e]", strings, value = TRUE)
grep("ab[^c]", strings, value = TRUE)
grep("^ab", strings, value = TRUE)
grep("\\^ab", strings, value = TRUE)
grep("abc|abd", strings, value = TRUE)
gsub("(ab) 12", "\\1 34", strings)
```

#Exercise 3
Write code to find countries in Gapminder with letter i or t, and ends with land, and replace land with LAND using backreference.
```{r}
grep("[it]", levels(gDat$country), value = TRUE) #contains i and t
grep("land$", levels(gDat$country), value = TRUE) # ends with land

#backreference to replace land with LAND
grep("land[it]", levels(gDat$country), value = TRUE)
grep("[it]|[land]", levels(gDat$country), value = TRUE)
grep("[it]|land$|\\$1LAND", levels(gDat$country), value = TRUE)

countries <- gsub("(.*[it].*)land$", "\\1LAND", levels(gDat$country)) 
grep("LAND", countries, value = TRUE)
```


# Part 4

Character classes 

[:digit:] or \d: digits, 0 1 2 3 4 5 6 7 8 9, equivalent to [0-9].
\D: non-digits, equivalent to [^0-9].
[:lower:]: lower-case letters, equivalent to [a-z].
[:upper:]: upper-case letters, equivalent to [A-Z].
[:alpha:]: alphabetic characters, equivalent to [[:lower:][:upper:]] or [A-z].
[:alnum:]: alphanumeric characters, equivalent to [[:alpha:][:digit:]] or [A-z0-9].
\w: word characters, equivalent to [[:alnum:]_] or [A-z0-9_].
\W: not word, equivalent to [^A-z0-9_].
[:xdigit:]: hexadecimal digits (base 16), 0 1 2 3 4 5 6 7 8 9 A B C D E F a b c d e f, equivalent to [0-9A-Fa-f].
[:blank:]: blank characters, i.e. space and tab.
[:space:]: space characters: tab, newline, vertical tab, form feed, carriage return, space.
\s: space, .
\S: not space.
[:punct:]: punctuation characters, ! " # $ % & ' ( ) * + , - . / : ; < = > ? @ [ \ ] ^ _ ` { | } ~.
[:graph:]: graphical (human readable) characters: equivalent to [[:alnum:][:punct:]].
[:print:]: printable characters, equivalent to [[:alnum:][:punct:]\\s].
[:cntrl:]: control characters, like \n or \r, [\x00-\x1F\x7F].
Note:
[:...:] has to be used inside square brackets, e.g. [[:digit:]].
\ itself is a special character that needs escape, e.g. \\d. Do not confuse these regular expressions with R escape sequences such as \t.

General modes for patterns 
POSIX (defaut) or Perl 
```{r}
(strings <- c("Axbc", "A.bc"))
pattern <- "A.b"
grep(pattern, strings, value = TRUE)
grep(pattern, strings, value = TRUE, fixed = TRUE)
```

```{r}
pattern <- "a.b"
grep(pattern, strings, value = TRUE)
grep(pattern, strings, value = TRUE, ignore.case = TRUE)
```


#Exercise 4
find continents in Gapminder with letter o in it
```{r}
grep("o{1}", levels(gDat$continent), value = TRUE, ignore.case = T)
```



