---
title: "Lecture6"
author: "Natalie Benoit"
date: "3/2/2021"
output: html_document
---
# Set up
```{r setup, include=FALSE}
library(gapminder)
library(tidyverse)
library(dplyr)
library(stringr)
gDat <- gapminder::gapminder
str(gDat)
```


# Part 1
```{r}
grep('\'', levels(gDat$country), value = TRUE)
```

\': single quote. You don't need to escape single quote inside a double-quoted string, so we can also use "'" in the previous example.
\": double quote. Similarly, double quotes can be used inside a single-quoted string, i.e. '"'.
\n: newline.
\r: carriage return.
\t: tab character.

cat() and print() to handle escape sequences differently, if you want to print a string out with these sequences interpreted, use cat().
```{r}
print("a\nb")
cat("a\nb")
```

Quantifiers
Quantifiers specify how many repetitions of the pattern.

*: matches at least 0 times.
+: matches at least 1 times.
?: matches at most 1 times.
{n}: matches exactly n times.
{n,}: matches at least n times.
{n,m}: matches between n and m times.

```{r}
(strings <- c("a", "ab", "acb", "accb", "acccb", "accccb"))
grep("ac*b", strings, value = TRUE)
grep("ac+b", strings, value = TRUE)
grep("ac?b", strings, value = TRUE)
grep("ac{2}b", strings, value = TRUE)
grep("ac{2,}b", strings, value = TRUE)
grep("ac{2,3}b", strings, value = TRUE)
```


# Exercise 1: 
line of code to find all countried with ee
```{r}
grep("e{2}", levels(gDat$country), value = TRUE)
```

# Part 2

Position of pattern within the string
^: matches the start of the string.
$: matches the end of the string.
\b: matches the empty string at either edge of a word. Don't confuse it with ^ $ which marks the edge of a string.
\B: matches the empty string provided it is not at an edge of a word.

```{r}
(strings <- c("abcd", "cdab", "cabd", "c abd"))
grep("ab", strings, value = TRUE)
grep("^ab", strings, value = TRUE)
grep("ab$", strings, value = TRUE)
grep("\\bab", strings, value = TRUE)
```

# Exercise 2:
write a code to find all files in main folder of repository
```{r}
path_towd <- getwd() # to get working directory
dir(path_towd) #see file names
dir(path_towd, pattern = ".R") # select for .R only -> does include .Rproj. Could I also do #.Rmd? 
```

