---
title: "leture7"
author: "Natalie Benoit"
date: "3/4/2021"
output: html_document
---

Install
```{r}
library(dplyr)
library(tidyverse)
library(praise)
```

Function basics
```{r}
#this won't do anything
function_name <- function(inputs) {
  output_value <- do_something(inputs)
  return(output_value)
}

#creating a function wont run it
calc_shrub_vol <- function(length, width, height) {
  volume <- length * width * height
  return(volume)
}

#call on arguments 
calc_shrub_vol(0.8, 1.6, 2.0)
# and store 
shrub_vol <- calc_shrub_vol(0.8, 1.6, 2.0)
```


## Exercise 1
Write a function that converts pounds to grams (there are 453.592 grams in one pound). It should take a value in pounds as the input and return the equivalent value in grams (i.e., the number of pounds times 453.592). Use that function to calculate how many grams there are in 3.75 pounds.
```{r}
pound_to_gram <- function(pound) {
  weight_g <- pound * 453.592
  return(weight_g)
}

pound_to_gram(3.75)
```


## Exercise 2
In your file lecture7-exercises.R, add a comment to this function so that you know what it does.
Use this function to print out the mass of a Spinosaurus that is 16 m long based on its reassembled skeleton.
Create a new version of this function called get_mass_from_length() that estimates the mass of an organism in kg based on its length in meters by taking length, a, and b as parameters. This lets us pass the function all 3 values that it needs to estimate a mass as parameters, which makes it much easier to reuse for all of the non-theropod species. Use this new function to estimate the mass of a Sauropoda (a = 214.44, b = 1.46) that is 26 m long.
```{r}
#estimating the mass (kg) of an organism in kg based on its length in meters
get_mass_from_length_theropoda <- function(length){
  mass <- 0.73 * length ^ 3.63
  return(mass)
}

get_mass_from_length_theropoda(16)

#new version
get_mass_from_length <- function(length,a, b){
  mass_ab <-length * 214.44 * 1.46
  return(mass_ab)
}

get_mass_from_length(26)
```

Default arguments
Defaults can be set for common inputs
```{r}
calc_shrub_vol <- function(length = 1, width = 1, height = 1) {
  volume <- length * width * height
  return(volume)
}

calc_shrub_vol()
calc_shrub_vol(width = 2)
calc_shrub_vol(0.8, 1.6, 2.0)
calc_shrub_vol(height = 2.0, length = 0.8, width = 1.6)
```


## Exercise 3
Use this function to estimate the mass of a Sauropoda (a = 214.44, b = 1.46) that is 22 m long (by setting a and b when calling the function).
Use this function to estimate the mass of a dinosaur from an unknown taxonomic group that is 16m long (by not setting a and b so that the default values are used).
```{r}
get_mass_from_length_defaults <- function(length, a = 39.9, b = 2.6){
  mass_default <-length * 214.44 * 1.46
  return(mass_default)
}
get_mass_from_length(22)

get_mass_from_length_defaults <- function(length, a = 39.9, b = 2.6){
  mass_default <-length * a * b
  return(mass_default)
}
get_mass_from_length(16)

#It is more useful passing them than having them fixed so they can be applied to many taxonomic groups. 
```


Named vs unamed 
```{r}
calc_shrub_vol(length = 2.0, width = 0.8, height = 1.6)
calc_shrub_vol(2.0, 0.8, 1.6)
```


Combining functions
```{r}
est_shrub_mass <- function(volume){
  mass <- 2.65 * volume^0.9
}

shrub_volume <- calc_shrub_vol(0.8, 1.6, 2.0)
shrub_mass <- est_shrub_mass(shrub_volume)

# piping 

shrub_mass <- calc_shrub_vol(0.8, 1.6, 2.0) %>%
  est_shrub_mass()
```


## Exercise 4
Write a function that converts kilograms into pounds (there are 2.205 pounds in a kilogram). Use that function along with your get_mass_from_length() function from Exercise 2 to estimate the weight, in pounds, of a 12 m long Stegosaurus. In Stegosauria, a has been estimated as 10.95 and b has been estimated as 2.64.
```{r}
kg_to_lbs <- function(lb){
  weight_lb <- lb / 2.205
  return(weight_lb)
}

get_mass_from_length <- function(length, a, b){
  mass_ab <-length * 10.95 * 2.64
  return(mass_ab)
}

stego_equ <- get_mass_from_length(10.95)
kg_to_lbs(stego_equ)

Stego_mass <- get_mass_from_length(10.95) %>%
  kg_to_lbs()

```











