---
title: "Lecture8"
author: "Natalie Benoit"
date: "3/9/2021"
output: html_document
---

```{r}
library(dplyr)
library(tidyverse)
```

Conditons and iterations
Operators:
==, !=
<, >
<=, >=
%in%


## Exercice 1
```{r}
w <- 10.2
x <- 1.3
y <- 2.8
z <- 17.5
colors <- c("red", "blue", "green")
masses <- c(45.2, 36.1, 27.8, 81.6, 42.4)
dna1 <- "attattaggaccaca"
dna2 <- "attattaggaacaca"

#w is greater than 10
w > 10
color = "green"
x > y
masses > 40
```



If statements 
Conditional statements generate logical values to filter inputs.
if statements use conditional statements to control flow of the program.
```{r}
x <- 6
if (x > 5){
  x <- x^2
}
x


x = 4
if (x > 5){
  x <- x^2
}
x

veg_type <- "tree"
volume <- 16.08
if (veg_type == "tree") {
  mass <- 2.65 * volume^0.9
  }
mass

```


#Exercise 2
```{r}
age_class = "sapling"
if (age_class == "sapling"){
  y <- 10
}
y
```



Alternative conditioning 
Often want to chose one of several options
Can add more conditions and associated actions with else if
```{r}
veg_type <- "grass"
volume <- 16.08
if (veg_type == "tree") {
  mass <- 2.65 * volume^0.9
} else if (veg_type == "grass") {
  mass <- 0.65 * volume^1.2
}
mass
```


## Excercise 3
```{r}
age_class <- "seedling"
if(age_class == "seedling"){
  y <- 5
}
y


age_class = "sapling"
if (age_class == "sapling"){
  y <- 10
}
y

age_class = "other"
if (age_class == "other"){
  y <- 0
}
y
```


Multiple if else & else if
Multiple ifs check each conditional separately
Executes code of all conditions that are TRUE
```{r}
x <- 5
if (x > 2){
  x * 2
}
if (x > 4){
  x * 4
}

x <- 5
if (x > 2){
  x * 2
} else if (x > 4){
  x * 4
}
```


Wrap conditions within a function 
```{r}
est_mass <- function(volume, veg_type){
  if (veg_type == "tree") {
    mass <- 2.65 * volume^0.9
  } else if (veg_type == "grass") {
    mass <- 0.65 * volume^1.2
  } else {
    print("I don't know how to convert volume to mass for that vegetation type")
    mass <- NA
  }
  return(mass)
}

est_mass(1.6, "tree")
est_mass(1.6, "grass")
est_mass(1.6, "shrub")
```


## Exercise 4
3 chunks of if else and or else (if nothing else)
```{r}
get_mass_from_length_by_name<-function(name, length) {
  if (name == "Stegosauria") {
    stego_mass <- 10.95*length^2.64
    return(stego_mass)
     
  } else if (name == "Theropoda"){
    thero_mass <- 0.73*length^3.63
    return(thero_mass)
    
  } else if (name == "Sauropoda"){
    sauro_mass <- 214.44*length^1.46
    return(sauro_mass)
    
  } else {
    na_mass <- NA*length^NA
    return(na_mass)
  }
}

get_mass_from_length_by_name("Stegosauria", 10)
get_mass_from_length_by_name("Theropoda", 8)
get_mass_from_length_by_name("Sauropoda", 12)
get_mass_from_length_by_name("Ankylosauria", 13)
```


## Exercise 5
```{r}
#modify to print 2-16
for (i in 2:16){
  print(i)
}

#modify to be multiplied by 3
for (i in 1:5 *3){
  print(i)
}

#complete code so that it prints out the name of each bird 
birds = c('robin', 'woodpecker', 'blue jay', 'sparrow')
for (i in 1:length(birds)){
  print(birds[i])
}

#Complete the code below so that it stores one area for each radius
radius <- c(1.3, 2.1, 3.5)
areas <- vector(mode = "numeric", length = length(vector))
for (i in 1:length(vector)){
  areas[i] <- pi * radius[i] ^ 2
}
areas

# Complete the code below to calculate an area for each pair of lengths and widths, store the areas in a vector, and after they are all calculated print them out:
lengths = c(1.1, 2.2, 1.6)
widths = c(3.5, 2.4, 2.8)
areas <- vector(length = i)
for (i in 1:length(areas)) {
  areas[i] <- lengths[i] * widths[i]
}
areas

```

