---
title: "LOTR"
author: "Natalie Benoit"
date: "2/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Packages
```{r cars}
library(tidyverse)
```

read csv files 
```{r pressure, echo=FALSE}
fship <- read_csv(file.path("data", "The_Fellowship_Of_The_Ring.csv"))

ttow <- read_csv(file.path("data", "The_Two_Towers.csv"))

rking <- read_csv(file.path("data", "The_Return_Of_The_King.csv")) 
```


collect untidy data
```{r}
lotr_untidy <- bind_rows(fship, ttow, rking)
str(lotr_untidy)

lotr_untidy
```

Exercise 1 - using pivot longer
```{r}
write.csv(lotr_untidy, file = file.path("data", "lotr_untidy.csv"))

lotr_tidy <-
  gather(lotr_untidy, key = 'Gender', value = 'Words', Female, Male)
lotr_tidy

lotr_tidy <-
  pivot_longer(lotr_untidy, cols=c(Female, Male), names_to = 'Gender', values_to = 'Words')

lotr_tidy <- arrange(lotr_tidy, Gender)

write_csv(lotr_tidy, path = file.path("data", "lotr_tidy.csv"))
```
Or, using the new language of pivot_longer(), we use values_to instead of values and instead of a key we specify names_to. We now also need to explicity write cols_to to designate the columns 'Male' and 'Female'.



Exercise 2

collect untidy data
```{r}
gender_untidy <- bind_rows(Female, Male)
str(gender_untidy)

lotr_untidy
 write.csv(gender_untidy, file = file.path("data", "gender_untidy.csv"))

```

Write R code to compute the total number of words spoken by each race across the entire trilogy. Do it two ways:

Using the lotr_tidy data frame as input.
```{r}
lotr_tidy %>% 
  count(Gender, Race, wt = Words)

```

Using film-specific or gender-specific, untidy data frames as the input data.
```{r}
gender_untidy %>% 
  select("Elf", "Hobbit", "Man") %>%
  summarise_all(funs(sum))
   
```


Exercise 3 - pivot wider 
```{r}
lotr_tidy %>% 
  pivot_wider(id_cols = c(Film, Gender), names_from = Words)

```


